---
title: "HW2_ISYE6414_ashish_dhiman"
author: "Ashish Dhiman | adhiman9@gatech.edu"
date: "2022-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

# Part 1

## Read Data and Summary

```{bash}
head -5 ./6414\-HW2\-taxes.csv
```

Since the first line does not have data we should skip it, also both the data columns are in 10k USD scale.

```{r}
df_tax = read.table(file ="./6414-HW2-taxes.csv",skip=1, sep=",",header=TRUE)
head(df_tax)
dim(df_tax)
summary(df_tax)
```

```{r}
df_tax = df_tax[1:(nrow(df_tax)-2),] #Remove last two empty rows
dim(df_tax)
```

## Question 1: Scatter Plot

```{r}
title_i = "Sales Price vs Annual Taxes (both in 10k USD)"
plot(x=df_tax$Taxes, y=df_tax$Sale.Price, type ="p",main = title_i)
```

$\color{blue}{\text{From the above plot, a linear realationship between Sales Price and Taxes is apparent.}}$

$\color{blue}{\text{The strength of the linear realtionship can also be tested with corealtion between x and y.}}$

```{r}
print (paste("Corealtion on full data:",round(cor(df_tax$Taxes,df_tax$Sale.Price),2)))
```

$\color{blue}{\text{A corelation of 0.88 is pretty significant and supports strong linear realtionship between x and y.}}$

## Question 2: Fit SLR

```{r}
#Fit SLR
model0 <- lm(Sale.Price ~ Taxes, data = df_tax)
model0

#Superpositioning regression line on 
ggplot(df_tax, aes(Taxes, Sale.Price)) + #aes(x,y)
  geom_point() +
  stat_smooth(method = lm, se = FALSE)
```

$$
From\ above, we\ have:\\
$$ $$
\hat{\beta}_0:= Intecept = 13.320,\\
$$ $$
\hat{\beta}_1:=Slope = 3.324,\ and\\
$$ $$
\hat{y} = \hat{\beta}_0 +\hat{\beta}_1 * x = 13.320 + 3.324 *x
$$

## Question 3: Meaning of beta1 ($\hat{\beta}_1$):

$\hat{\beta}_1$ implies the slope of the SLR line we have fit to the data. In other words, it tells us, the change recorded in y (on average) for every one unit of change in x.

$\color{blue}{\text{In this case, this implies for every 10k USD change in Taxes, the Price goies up by 3,324 USD (on average).}}$

## Question 4: Meaning of beta0 ($\hat{\beta}_0$):

$\hat{\beta}_0$ implies the predicted value of y given x is zero.

$\color{blue}{\text{In this case, this implies for 0 USD in taxes (let's assume), the the expected prices is 13,320 USD}}.$

$\color{blue}{\text{This case is improbable but still possible, and would imply that the state has relaxed the tax rate to 0%}}.$

This could be done by state to foster investment my individuals and firms.

## Question 5: value of s, $s^2$ and SSE:

s and s\^2

```{r}
s_squared = sum(sapply(model0["residuals"], function(x) x^2))/(nrow(df_tax)-2)
s_squared
s_squared^0.5
```

### SSE from fitted values

$SSE = \sum_{i \in all\ data\ points}([y_i - \hat{y}_i]^2)$

```{r}
y_hat = fitted(model0)
y_act = df_tax$Sale.Price

sse = sum((y_act - y_hat)^2)
print (paste("sse is",sse))

sse/(nrow(df_tax)-2)
```

# Part 2

## Question 6: Least Square Estimate of beta0, beta1:

$$
\hat{\beta}_1 = \frac{(\sum_i{x_iy_i}) - n\overline{x}\overline{y}}{\sum_i{{x_i}^2} - n\overline{x}^2},\\$$ $$
we\ have:\\
$$ $$
\sum_i{x_iy_i} = 1697.8,\\$$ $$
n\overline{x}\overline{y} = n * \frac{\sum_i{x_i}}{n} * \frac{\sum_i{y_i}}{n},\\$$ $$
\sum_i{{x_i}^2} = 157.42,\\$$ $$
n\overline{x}^2 = n * \frac{\sum_i{x_i}}{n} = \sum_i{x_i} = 14*(43/14)^2
$$

```{r}
num = 1697.8 - (14 * (43/14) * (572/14))
num
denom = 157.42 - (14*(43/14)^2)
denom
beta1 = num/denom
print (paste("Intecept (or beta0)",beta1))
```

$$
\hat{\beta}_0 = \overline{y} - \hat{\beta}_1\overline{x}
$$

```{r}
beta0 = (572/14) - beta1*(43/14)
print (paste("Intecept (or beta0)",beta0))
```

## Question 7: Calculate SSE

$$
SSE = SS_{yy} - \hat{\beta}_1SS_{xy},$$ $$
SS_{yy} = \sum_i{(y_i - \overline{y})^2} = \sum_i{y_i}^2 - n\overline{y}^2$$ $$
SS_{xy} = (\sum_i{x_iy_i}) - n\overline{x}\overline{y}
$$

```{r}
SS_yy = 23530 - 14 *((572/14))^2
SS_xy = num #from ques7
SSE = SS_yy - beta1 * SS_xy
sigma2 = SSE/(14-2)
print (paste("SSE is",SSE))
print (paste("Sigma squared is",sigma2))
```

## Question 8: y for x =3.7

```{r}
y_act = 46.1
y_hat = beta0 + beta1 * 3.7
residual = y_act - y_hat
print (paste("Predicted Y value is",y_hat))
print (paste("Corresponding Residual is",residual))
```
